Fast Sorting on Modern Computers

### 2005-08-02 docs/tech_report/heap.tex
fixed up the references

In the graph, the base heapsort's results flat-line for a step longer than the
other heapsorts' results. This is due to being in-place, since twice as many
keys fit in the cache without a corresponding increase in misses. For the same
reason, the base heapsorts have no compulsory misses\footnote{See Section
\ref{TODO}.}, while the out-of-place sorts have one compulsory miss in eight,
when the auxiliary heap is being populated.

@@

Figure \ref{base and cache heapsorts-1}(a) shows the behaviour of each
branch in the base heapsort algorithm. The ``creation, has children'' branch
is from line \textbf{xxxxx} of Figure \ref{Simple Base Heapsort}.

@@

>>>>

In the graph, the base heapsort's results flat-line for a step longer than the
other heapsorts' results. This is due to being in-place, since twice as many
keys fit in the cache without a corresponding increase in misses. For the same
reason, the base heapsorts have no compulsory misses, while the out-of-place
sorts have one compulsory miss in eight, when the auxiliary heap is being
populated.

@@

Figure \ref{base and memory-tuned heapsorts-1}(a) shows the behaviour of each
branch in the base heapsort algorithm. The ``creation, has children'' branch
is from line \textbf{xxxxx} of Figure \ref{Simple Base Heapsort}.

@@

%TODO read over the merged bits. might not be quite right

### 2005-08-02 docs/tech_report/quick.tex
fixed up the references

We use several variations of quicksort to test this theory. \ref{bently, i
ithkn} suggests partitioning using a median-of-9 instead of a median-of-3. We
created sorts using no median, a median-of-3, and pseudo-medians-of-5, -7 and
-9. The pseudo-medians used repeated medians-of-3 instead of a real
median, as suggestion by \ref{bently, again, i think}. 
%use a cut off

@@

The results of these tests are shown in Figure \ref{Branch prediction
performance for base quicksort without a median, with median of 3, and with
pseudo-medians of 5, 7 and 9}. There is little variance in insertion sort misses
between the sorts using pseudo-medians, but the median-of-3 quicksort has
6\% fewer branches, and a corresponsing decrease in mispredictions.
%TODO to whom do we owe the idea of a pseudo median? Bently/McIlroy?
%TODO when does it cut out, and why did we choose tere (benlty?)

@@

>>>>

We use several variations of quicksort to test this theory.
\cite{BentleyMcIlroy93} suggests partitioning using a median-of-9 instead of a
median-of-3. We created sorts using no median, a median-of-3, and
pseudo-medians-of-5, -7 and -9. As suggested, our pseudo-medians used repeated
medians-of-3 instead of a real median, in order to reduce the cost of their
calculation. It was also suggested that once the partition sized were small
enough, there was no more advantage to be gained by using a median of greater
than three, and so a median-of-3 was used below a threshhold of 40.

@@

The results of these tests are shown in Figure \ref{Branch prediction
performance for base quicksort without a median, with median of 3, and with
pseudo-medians of 5, 7 and 9}. There is little variance in insertion sort misses
between the sorts using pseudo-medians, but the median-of-3 quicksort has
6\% fewer branches, and a corresponsing decrease in mispredictions.

@@

%TODO do we add more graphs here?

### 2005-08-02 docs/tech_report/radix.tex
fixed up the references

\section{Memory-tuned Radixsort}
Many improvements exist to radixsort, and some are suggested by Sedgewick and
LaMarca. Some of these improvements help to lower cache misses and instruction
count, both of which slow down radixsort significantly. LaMarca suggests an
improvement by Friend\footnote{See \ref{TODO friend[8]}.} that while the current
byte is being put in place, the counting of the next byte can be done. This
increases the memory requirements, but this is not significant.

>>>>

\section{Memory-tuned Radixsort}
Many improvements exist to radixsort, and some are suggested by Sedgewick and
LaMarca. Some of these improvements help to lower cache misses and instruction
count, both of which slow down radixsort significantly. LaMarca suggests an
improvement by Friend\footnote{See \cite{Friend56:152}.} that while the current
byte is being put in place, the counting of the next byte can be done. This
increases the memory requirements, but this is not significant.
